---
title: "Project Part 3" 
author: "Yama Baluch"
fontsize: 12pt
geometry: margin=1in
urlcolor: black
output: pdf_document
header-includes:
- \usepackage{setspace}
- \onehalfspacing
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, highlight=FALSE)
```


```{r, echo=FALSE, include=FALSE}
# For Part 1, use this chunk for reading in the data with a path that identifies you.
library(readr)
shootings <- read_csv("shootings.csv")
View(shootings)

```
## Question
Does the implementation of Body Cameras make a significant difference in the likelihood 
of police brutality cases?

## Data description
A data set of all police killings from January 1st, 2015 to June 15th, 2020 will
be used to answer this question. The data set is a population and was published
by the Washington Post and is being constantly updated. The data set has 15 variables
to describe and identify the victim and the incident. The two variables that will be of
interest for this question are "Armed" and "body_camera," which identifies the type of
weapon that the victim was carrying and if the officer involved with the incident was
wearing a body camera. Since the question is about police brutality, we will focus on
the victims whose Armed category was identified as "unarmed."  
#Reference 1    

## Generalization
We will think of the population of unarmed victims as a sample of an even larger
population. Looking at the population as a sample of a super population, the outcomes
of the test statistic can be generalized to the population of police brutality victims
from the years 2021 to 2023. Beyond that, their is no telling how the future of body
camera usage will be like. If more funding is put into police body camera programs, the
data will change drastically and it is possible that more incidents will involve body
cameras due to them becoming more prominent.

## Test selection
A left-sided one sample z-test for proportions will be used to answer the question. This
is the most appropriate test because it is used to test characteristics of a population.
The characteristic being tested in this case would be whether or not the officers were
wearing a body camera. For this test, we will think of the population of unarmed victims
as the sample. The population that our sample will be representing will be the super
population and will include unarmed victims in the years from 2021 to 2023. In terms of
assumption checking, there were at least 5 successes and at least 5 failures, so the data
set meets the requirements of the test. A success in this case would be defined as "True"
the officer was wearing a body camera, while a failure would be defined as "False" they
were not wearing a body camera.  
The null hypothesis for this test will be that, on average, the proportion of police brutality
incidents that involved police that wore body cameras is 50%. The alternative hypothesis will
be that, on average, the proportion of police brutality incidents that involved police that wore
body cameras is less than 50%.
```{r}
Unarmed<-shootings[ which(shootings$armed=="unarmed"), ]
table(Unarmed$body_camera)
z.prop<-prop.test(55,348,p=0.50,alternative="less", correct=FALSE)
z.prop
phat <- 55/348
z <- (phat-0.5)/sqrt(0.5*0.5/348)
z
p.value <- pnorm(z)
p.value
sqrt(z.prop$statistic)
z.prop$p.value
```
## Conclusion
From the 348 unarmed victims, 55 were killed by officers wearing body cameras. The
proportion of these incidents were 15.87%. The proportion of incidents that involved
police with body cameras was less than the proportion of incidents that did not involve
body cameras, which gives us reasonable evidence that the implementation of body cameras
make a significant difference in the likelihood of police brutality cases.  
Since all recorded police brutality incidents from the years 2015 to 2020 were taken into account, I would
say that our findings were very representative of the effects of officers wearing body cameras.
We could generalize that from the years 2021 to 2023, around 16% of all police brutality incidents would
involve an officer wearing a body camera. There could be drastic changes to police departments in the years
beyond that. If more police departments implement body cameras, the proportion of incidents with body cameras
would likely increase due to there being more officers wearing body cameras than officers not wearing body
cameras.


## References
1. <https://www.kaggle.com/ahsen1330/us-police-shootings>.
